---
title: "HW2 Data Wrangling I"
author: "Minghui Wang" 
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```
[HW2](https://p8105.com/homework_2.html) assignment reinforces ideas in [Data Wrangling I](https://p8105.com/topic_data_wrangling_i.html). Here are the codes focusing on question 1-3.

# Problem 1
## Import and clean the NYC Transit CSV
```{r}
trans_entr_df <- 
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  col_types = cols( "Route8" = "c","Route9" = "c","Route10" = "c","Route11" = "c")) |> 
  janitor::clean_names() |> 
  select(
    line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada
  ) |> 
  mutate(entry = if_else(entry == "Yes", TRUE, FALSE))
```

```{r}
dim(trans_entr_df)
```
A short paragraph about this dataset:
This dataset contains 32 variables, with 22 being character variables (such as Division, Line, Station Name, and Route1 to Route7), 8 being numeric variables (including Station Latitude, Station Longitude, Entrance Latitude, and Entrance Longitude), and 2 being logical variables (ADA and Free Crossover). So far, I have standardized the column names by converting them to lowercase and treated all route variables as character types. The resulting dataset has dimensions of 1868 rows and 19 columns. Currently, the data is not tidy because the route numbers are spread across multiple columns (from "Route1" to "Route11") rather than being combined into a single column as a variable.

# Problem 2
```{r}

```

```{r}

```
# Problem 3
```{r}

```

